Full Brain BOLD Signal Parameter Estimation \\ using Particle Filters

### 2010-08-29 doc/presentation/introduction.tex
notes are in and ready to print

\section{Introduction}
\begin{frame}{Balloon Model}
\begin{figure}
\includegraphics[width=10cm]{model2}
\caption{
    \tiny
    \cite{Riera2003}
}
\end{figure}
\note{Its been known for about 20 years that changes in Blood Oxygen Level alter
        T2$^*$ relaxation times. The basis of contrast in fMRI is 
        changes in the so-called spin-spin relaxation (or T2$^*$ relaxation time
        of tissue). The rate of loss in nuclear spin coherence depends on the molecules in the voxel being
        measured \emph{and the surrounding voxels}. For this reason the signal 
        detected from BLOOD doesn't just depend on the amount of oxygen, but also the
        amount of De-oxygenated hemoglobin. The question then, is what drives changes
        in De-oxygenated Hemoglobin and blood volume?

        Well the brain has no local oxygen storage, so it is completely dependent on 
        fresh Oxygenated-Hemoglobin. When the brain becomes more active, firing rates
        for neurons increase by several orders of magnitude, and recharging these neurons
        requires additional oxygen for matabolism. }
\note{Rates from below 1 Hz to 5 to 6 Hundred Hz}
\note{Change in average firing rates increase metabolism and sympathetic response causes
        increased inflow of oxygenated blood.}
\note{In the version of the Balloon model I used, the flow is locked to metabolism.}
\note{According the balloon model this increases the local blood volume, and the outflow pressure.}
\note{So the signal is altered by changes in the ratio of Deoxyhemoglobin and Oxygenated Hemoglobin,
        which are a function of Volume and Oxygenation.}
\end{frame}

\begin{frame}{Activation Chain}
\begin{figure}
    \includegraphics[width=\textwidth]{graphs}
\caption{
    \tiny
    \cite{Buxton2004}
}
\end{figure}
\end{frame}

@@

  \begin{column}{.3\textwidth}
    \begin{itemize}
        \item State Variables:
        $${\color{mblue}s(t)}, {\color{mblue}f(t)}, {\color{mblue}v(t)}, {\color{mblue}q(t)}$$
        \item Parameters:
        $${\color{mred}\epsilon}, {\color{mred}\tau_s}, {\color{mred}\tau_f}, {\color{mred}\alpha}, 
                    {\color{mred}\tau_0}, {\color{mred}E_0}, {\color{mred}V_0}$$
        \item Constants:
        $$a_1, a_2$$
        \item Input:
        $$\color{mgreen}u(t)$$
    \end{itemize}
  \end{column}
  \end{columns}
  \note{System is dissipative:\\
  If no input is provided, it will rather quickly decay.\\
  Constant input leads to steady state response.\\
  First term in $\dot{q}$ is the metabolism term.\\
  Outflow is the $v^{1/\alpha}$ term\\
  Significant nonlinearity.}
\end{frame}

>>>>

\section{Introduction}
\begin{frame}{Balloon Model}
\begin{figure}
\includegraphics[width=10cm]{model2}
\caption{
    \tiny
    \cite{Riera2003}
}
\end{figure}
\note{
    \begin{itemize}
        \item FMRI has been around for about 20 years
        \item 10 Years Since the First BOLD Model
        \item Basis of FMRI, T2$^*$ relaxation time, blood oxygen levels
        \item Windkessel/Balloon Model, Volume Changes driving fluctations
            in blood oxygen 
        \item Brain has no local oxygen storage, dependent on blood
        \item brain active, firing rates of neurons increase 10 or 100x
        \item recharging requires additional oxygen ... metabolism. 
        \item Increased metabolism drives increased blood flow
        \item Increased blood flow increases volume

        \item Rates from below 1 Hz to 5 to 6 Hundred Hz
    \end{itemize}
}
\end{frame}

\begin{frame}{Activation Chain}
\begin{figure}
    \includegraphics[width=\textwidth]{graphs}
\caption{
    \tiny
    \cite{Buxton2004}
}
\end{figure}
\note{
    \begin{itemize}
        \item CMRO2 Increased Metabolism Causes Initial Drop In Oxygen
        \item CBF Flow Quickly Compensates Increases Oxygen
        \item CBV Increases, Decides the outflow, local oxygen content
        \item This delays result in the characteristic BOLD response
        \item CBF CMRO2 Locked 
    \end{itemize}
}
\end{frame}

@@

  \begin{column}{.3\textwidth}
    \begin{itemize}
        \item State Variables:
        $${\color{mblue}s(t)}, {\color{mblue}f(t)}, {\color{mblue}v(t)}, {\color{mblue}q(t)}$$
        \item Parameters:
        $${\color{mred}\epsilon}, {\color{mred}\tau_s}, {\color{mred}\tau_f}, {\color{mred}\alpha}, 
                    {\color{mred}\tau_0}, {\color{mred}E_0}, {\color{mred}V_0}$$
        \item Constants:
        $$a_1, a_2$$
        \item Input:
        $$\color{mgreen}u(t)$$
    \end{itemize}
  \end{column}
  \end{columns}
  \note{
    \tiny
    \begin{itemize}
    \item System is dissipative,
    \item Constant input leads to steady state 
    \item $u(t)$ is the input
    \item States
    \begin{itemize}
    \item s - Neural Response (Flow Inducing Signal)
    \item f - Inflow of Oxygenated Blood
    \item v - Local Blood Volume
    \item q - Deoxygenated hemoglobin content
    \end{itemize}
    \item Parameters
    \begin{itemize}
    \item $\epsilon$ - efficiency
    \item $\tau_s$ - Falloff time for Flow Inducing Signal
    \item $\tau_f$ - Onset time for Flow Inducing Signal
    \item $\alpha$ - Grubbs Constant determining outflow volume
    \item $\tau_0$ - Time Constant for volume, deoxygenated hemoglobin
    \item $E_0$ - Resting Oxygen Metabolism
    \item $V_0$ - Resting Metabolism
    \end{itemize}
    \item $a_1$,$a_2$ are functions of the imaging modality,
            and to a vary somewhat with $E_0$. 
    \item $a_1 = k_1*E_0 + k_2*E_0$
    \item $a_2 = k_2*E_0 + k_3$
    \end{itemize}
  }
\end{frame}

### 2010-08-29 doc/presentation/main.tex
notes are in and ready to print

\usepackage{pgfpages}
%\setbeameroption{show notes on second screen}

@@

%0
\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
  \note{Introduction describes FMRI and the Balloon Model}
  \note{Potential Approaches - describes previous used methods of analysis}
  \note{Particle Filter - derives and discusses the BOLD model}
  \note{Methods - The methods used in this work, how to recreate.}
\end{frame}

@@

\begin{frame}[allowframebreaks]
  \bibliographystyle{plain}
  \bibliography{../library}
\end{frame}

\appendix
\begin{frame}{SPM vs. Mutual Information Map, SPM}
\begin{figure}
\setcounter{subfigure}{0}
\subfigure[SPM Results]{\includegraphics[width=.8\textwidth]{images/spm_hm}}
\subfigure{\includegraphics[scale=.5]{images/scale1}}
\note{SPM results. Units of activation are in Student's $t$-scores; higher 
        indicates higher assurance that the signal cannot have occurred 
        through noise alone.}
\end{figure}
\end{frame}

@@

>>>>

\usepackage{pgfpages}
%\setbeameroption{show notes on second screen}
\setbeameroption{show notes}

@@

%0
\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
  \note{
  \begin{itemize}
  \item Introduction - FMRI and the Balloon Model
  \item Nonlinear Regression - overview prior works and the particle filter
  \item Methods - how the particle was configured, and the preprocessing
  \item Results - Several different types of simulation and real data 
  \item Conclusion - Why this is important
  \end{itemize}
  }
\end{frame}

@@

\appendix

\section{Backup}
\begin{frame}{Balloon Flowchart}
\begin{figure}
\includegraphics[width=\textwidth]{backup_balloon}
\end{figure}
\end{frame}

\begin{frame}{SPM vs. Mutual Information Map, SPM}
\begin{figure}
\setcounter{subfigure}{0}
\subfigure[SPM Results]{\includegraphics[width=.8\textwidth]{images/spm_hm}}
\subfigure{\includegraphics[scale=.5]{images/scale1}}
\note{SPM results. Units of activation are in Student's $t$-scores; higher 
        indicates higher assurance that the signal cannot have occurred 
        through noise alone.}
\end{figure}
\end{frame}

@@

\begin{frame}{Particle Filter Results: Histogram}
\begin{figure}
\centering
\includegraphics[clip=truew,trim=8cm 4cm 8cm 4cm,width=.8\textwidth]{realhist}
\note{Histogram of parameters in active regions ($M.I. > .15$).}
\end{figure}
\end{frame}

\section{Bibliography}
\begin{frame}[allowframebreaks]
  \bibliographystyle{plain}
  \bibliography{../library}
\end{frame}
